\chapter{Conclusion}

\section{Summary}

In this diploma thesis we were able to create a generator of spatio-temporal and textual social media data. We used real points of interest and ratings and reviews 
for these points as those were crawled by real users of the online travel service TripAdvisor. The utilization of these real points and the directions obtained 
from Google Directions API resulted in the creation of realistic routes, check-ins and GPS traces. Also, the fact that the generator uses random factors in order 
to take decisions about the place, duration and number of daily visits enhances the realistic aspect of the generated data. Additionally, most of these factors 
are instantiated by the corresponding input parameters, so that the generator can produce datasets with characteristics that vary. Therefore, we were able to create 
users that resemble real users found in well-known online 
social networking services. 

The daily execution of the generator for a significant period of time resulted in the assemblance of a Big Data dataset of users with daily routes and 
check-ins to points of interest. More specifically, the generator created 9464 users, 1586537 check-ins and 38800019 GPS traces, which sum up to 3 GB data. 
In this way, we were able to re-create realistic social media data that we couldn't have access to due to privacy restrictions imposed by 
current social networking services. 

With these data available, we were able to evaluate the distributed storage system that HBase provides and is used by a significant 
number of such social networking services. For this reason, we stored the created dataset into an HBase cluster using an appropriate data model and by 
ensuring that the cluster will be data balanced. Then, 
we implemented several queries over the available data that are common to such 
services. The queries were implemented using HBase coprocessors, which move the computation of intermediate results and other complex calculations 
to the region servers that contain the respective data, decongesting the client from a heavy computational load. We, then, created a workload with these 
types of queries in order to test the scalability of the HBase cluster. We measured the latency and throughput of 
the HBase cluster while changing the number of concurrent queries and cluster size. We came to the conclusion that 
the HBase cluster which contains the social networking data produced by the generator and stored with the specific data model, is 
scalable as it has the expected behavior of common scalable systems. 
Therefore, HBase provides indeed good performance and storage for Big Data social media services. 

\section{Related Work}

There are already generators that produce social media data. One such generator is DATAGEN, which is a social network data generator, provided by LDBC-SNB 
(Linked Data Benchmark Council - Social Network Benchmark \cite{26}). The LDBC-SNB aims to provide a realistic workload, consisting of a social network and a set of queries. 
DATAGEN is a synthetic data generator responsible for generating datasets for the three LDBC-SNB workloads: the Interactive, the Business Intelligence and the Analytical. 
DATAGEN mimics features of those found in real social network. They create a synthetic dataset due to the fact that it is difficult to find data with all the 
scaling characteristics their benchmark requires and collecting real data can be expensive or simply not possible due to privacy concerns.

Also, there is LFR-Benchmark generator \cite{27} which provides the network structure of a social network, being able to mimic some of the properties of real 
social media. The LFR generator produces social network graphs with power-law degree distribution and with implanted communities within the network.


\section{Future Work}

There are several ways in which the current diploma thesis could be extended. As fas the generator is concerned, the produced data could be enhanced by also adding more 
textual data such as Facebook posts or Twitter tweets. Also, the generator could use a photo and video pool in order to simulate the upload of photos and videos, 
which is a basic functionality of most social networking services. As far as the generator execution is concerned, the generator could be instantiated with other 
values in order to create a more polymorphic dataset. Also, the usage of Google Maps API for Work enables 100000 directions requests per 24 hour period, which is 
significantly more than the limit of 2500 daily requests for the users of the free API. Therefore, the generator can be used more efficiently for users 
subscribed to the Google Work services. 

Moreover, regarding the implemented queries, there are many more that can be implemented. Also, there could be queries over the data of GPS traces which are 
created by the generator. In this way, the created workload can be more efficient into the scalability testing of the HBase cluster. Finally, other 
distributed storage and management systems could be evaluated in a similar way, such as Hadoop. 
